---
title: "Method of compartment"
author: "MJY"
date: "2026-04-17"
output: html_document
---

################################################################################ 

# 

# OUTLINES:

# 1. Identify A/B compartments using multiple methods
# 2. Compare A compartments between methods and visualize overlap
# 3. Compare B compartments between methods and visualize overlap

# 

################################################################################ 

#' ## sample ccl4 18w
#' ## 1. Identify A/B compartments using multiple methods

```{r}
# Load required libraries
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(eulerr))

# =========================
# Load HiCExplorer compartment results
# =========================

hicex_files <-"F:/share/results/compartment/hicexplore/ccl4_18w_hicex_final.bedgraph"
hicex <- fread(hicex_files, header = TRUE, stringsAsFactors = FALSE)

# Remove rows containing zero values
hicex <-hicex %>%
  filter(rowSums(. == 0) == 0)

# Create unique region identifier (chr:start:end)
hicex$region_id <- apply(hicex, 1, function(x) paste(x[1], x[2], x[3], sep = ":"))


# =========================
# Load HiTC compartment results
# =========================

hitc_files <-"F:/share/results/compartment/hitc/ccl4_18w_pca.bedgraph"
hitc <- fread(hitc_files, header = TRUE, stringsAsFactors = FALSE)

# Remove rows containing zero values
hitc <-hitc %>%
  filter(rowSums(. == 0) == 0)

# Define a function to classify PC1 score into A/B compartments
classify_score <- function(score) {
  if (!is.na(score)) {
    if (score > 0) {
      return("A")
    } else {
      return("B")
    }
  } else {
    return(NA)  # Keep NA values
  }
}

# Apply classification function to score column
hitc$Compartment <- sapply(hitc$score, classify_score)

# Create unique region identifier
hitc$region_id <- apply(hitc, 1, function(x) paste(x[1], x[2], x[3], sep = ":"))


# =========================
# Load bin reference file
# =========================

# Read bin_200000_PC1.xlsx file to obtain all genomic regions
bin_file <- "F:/share/results/compartment/bedgraph/bin_200000_PC1.xlsx"
bins <- read_excel(bin_file)

# Create unique region identifier (chr:start:end)
bins$region_id <- apply(bins, 1, function(x) paste(x[1], x[2], x[3], sep = ":"))


# =========================
# Align compartment data across methods
# =========================

# Define a function to align and rename compartment data
align_and_rename <- function(bins, data, name) {
  aligned_data <- merge(bins, data, by = "region_id", all.x = TRUE)
  aligned_data <- aligned_data[, c("region_id", "Compartment")]
  colnames(aligned_data)[colnames(aligned_data) == "Compartment"] <- name
  return(aligned_data)
}

# Align data from different methods
hitc_aligned_data <- align_and_rename(bins, hitc, "hitc")
hicex_aligned_data <- align_and_rename(bins, hicex, "hicex")

# Merge all aligned datasets
aligned_data <- Reduce(function(x, y) merge(x, y, by = "region_id", all.x = TRUE), 
                       list(hitc_aligned_data, hicex_aligned_data))

# Remove rows containing NA values
aligned_data <- aligned_data[complete.cases(aligned_data), ]


# =========================
# Extract A/B compartments for each method
# =========================

# Define method names
methods <- c("hitc", "hicex")

# Extract indices of A compartments
compartment_a <- lapply(methods, function(method) which(aligned_data[[method]] == "A"))

# Extract indices of B compartments
compartment_b <- lapply(methods, function(method) which(aligned_data[[method]] == "B"))

# Assign names to output lists
names(compartment_a) <- methods
names(compartment_b) <- methods

```

#' ## 2. Identify A/B compartments using multiple methods

```{r}

# =========================
# Define compartment sets
# =========================

# Extract A compartments from different methods
A <- compartment_a$hitc
B <- compartment_a$hicex


# =========================
# Calculate overlap statistics
# =========================

# Define a function to count unique and overlapping regions
count_region <- function(set1, set2) {
  list(
    A_only = length(setdiff(set1, set2)),
    B_only = length(setdiff(set2, set1)),
    A_B_only = length(intersect(set1, set2))
  )
}

# Compute region counts
region_counts <- count_region(A, B)

# Print results
print(region_counts)


# =========================
# Visualize overlap using Euler diagram
# =========================

print("ccl4 18w compartment A")
plot(euler(compartment_a),
     fills = list(fill=c("#008080", "#ffbb44"),
                  alpha=0.8),
     quantities = TRUE,
     legend = list(labels=c("HiTC", "HiCExplorer"),
                   cex=1.2))


```

#' ## 3. Compare B compartments between methods and visualize overlap

```{r}
# =========================
# Define compartment sets
# =========================

# Extract B compartments from different methods
A <- compartment_b$hitc
B <- compartment_b$hicex


# =========================
# Calculate overlap statistics
# =========================

# Define a function to count unique and overlapping regions
count_region <- function(set1, set2) {
  list(
    A_only = length(setdiff(set1, set2)),
    B_only = length(setdiff(set2, set1)),
    A_B_only = length(intersect(set1, set2))
  )
}

# Compute region counts
region_counts <- count_region(A, B)

# Print results
print(region_counts)


# =========================
# Visualize overlap using Euler diagram
# =========================


# Basic Euler plot
print("ccl4 18w compartment B")
plot(euler(compartment_b),
     fills = list(fill=c("#008080", "#ffbb44"),
                  alpha=0.8),
     quantities = TRUE,
     legend = list(labels=c("HiTC", "HiCExplorer"),
                   cex=1.2))
```

#' ## sample con 18w
#' ## 1. Identify A/B compartments using multiple methods

```{r}
# =========================
# Load HiCExplorer compartment results
# =========================

hicex_files <-"F:/share/results/compartment/hicexplore/con_18w_hicex_final.bedgraph"
hicex <- fread(hicex_files, header = TRUE, stringsAsFactors = FALSE)

# Remove rows containing zero values
hicex <-hicex %>%
  filter(rowSums(. == 0) == 0)

# Create unique region identifier (chr:start:end)
hicex$region_id <- apply(hicex, 1, function(x) paste(x[1], x[2], x[3], sep = ":"))


# =========================
# Load HiTC compartment results
# =========================

hitc_files <-"F:/share/results/compartment/hitc/con_18w_pca.bedgraph"
hitc <- fread(hitc_files, header = TRUE, stringsAsFactors = FALSE)

# Remove rows containing zero values
hitc <-hitc %>%
  filter(rowSums(. == 0) == 0)

# Define a function to classify PC1 score into A/B compartments
classify_score <- function(score) {
  if (!is.na(score)) {
    if (score > 0) {
      return("A")
    } else {
      return("B")
    }
  } else {
    return(NA)  # Keep NA values
  }
}

# Apply classification function to score column
hitc$Compartment <- sapply(hitc$score, classify_score)

# Create unique region identifier
hitc$region_id <- apply(hitc, 1, function(x) paste(x[1], x[2], x[3], sep = ":"))


# =========================
# Load bin reference file
# =========================

# Read bin_200000_PC1.xlsx file to obtain all genomic regions
bin_file <- "F:/share/results/compartment/bedgraph/bin_200000_PC1.xlsx"
bins <- read_excel(bin_file)

# Create unique region identifier (chr:start:end)
bins$region_id <- apply(bins, 1, function(x) paste(x[1], x[2], x[3], sep = ":"))


# =========================
# Align compartment data across methods
# =========================

# Define a function to align and rename compartment data
align_and_rename <- function(bins, data, name) {
  aligned_data <- merge(bins, data, by = "region_id", all.x = TRUE)
  aligned_data <- aligned_data[, c("region_id", "Compartment")]
  colnames(aligned_data)[colnames(aligned_data) == "Compartment"] <- name
  return(aligned_data)
}

# Align data from different methods
hitc_aligned_data <- align_and_rename(bins, hitc, "hitc")
hicex_aligned_data <- align_and_rename(bins, hicex, "hicex")

# Merge all aligned datasets
aligned_data <- Reduce(function(x, y) merge(x, y, by = "region_id", all.x = TRUE), 
                       list(hitc_aligned_data, hicex_aligned_data))

# Remove rows containing NA values
aligned_data <- aligned_data[complete.cases(aligned_data), ]


# =========================
# Extract A/B compartments for each method
# =========================

# Define method names
methods <- c("hitc", "hicex")

# Extract indices of A compartments
compartment_a <- lapply(methods, function(method) which(aligned_data[[method]] == "A"))

# Extract indices of B compartments
compartment_b <- lapply(methods, function(method) which(aligned_data[[method]] == "B"))

# Assign names to output lists
names(compartment_a) <- methods
names(compartment_b) <- methods

```

#' ## 2. Identify A/B compartments using multiple methods

```{r}

# =========================
# Define compartment sets
# =========================

# Extract A compartments from different methods
A <- compartment_a$hitc
B <- compartment_a$hicex


# =========================
# Calculate overlap statistics
# =========================

# Define a function to count unique and overlapping regions
count_region <- function(set1, set2) {
  list(
    A_only = length(setdiff(set1, set2)),
    B_only = length(setdiff(set2, set1)),
    A_B_only = length(intersect(set1, set2))
  )
}

# Compute region counts
region_counts <- count_region(A, B)

# Print results
print(region_counts)


# =========================
# Visualize overlap using Euler diagram
# =========================

print("con 18w compartment A")
plot(euler(compartment_a),
     fills = list(fill=c("#008080", "#ffbb44"),
                  alpha=0.8),
     quantities = TRUE,
     legend = list(labels=c("HiTC", "HiCExplorer"),
                   cex=1.2))



```

#' ## 3. Compare B compartments between methods and visualize overlap

```{r}
# =========================
# Define compartment sets
# =========================

# Extract B compartments from different methods
A <- compartment_b$hitc
B <- compartment_b$hicex


# =========================
# Calculate overlap statistics
# =========================

# Define a function to count unique and overlapping regions
count_region <- function(set1, set2) {
  list(
    A_only = length(setdiff(set1, set2)),
    B_only = length(setdiff(set2, set1)),
    A_B_only = length(intersect(set1, set2))
  )
}

# Compute region counts
region_counts <- count_region(A, B)

# Print results
print(region_counts)


# =========================
# Visualize overlap using Euler diagram
# =========================

# Basic Euler plot
print("con 18w compartment B")
plot(euler(compartment_b),
     fills = list(fill=c("#008080", "#ffbb44"),
                  alpha=0.8),
     quantities = TRUE,
     legend = list(labels=c("HiTC", "HiCExplorer"),
                   cex=1.2))
```

