
################################################################################ 

# 

# OUTLINES:

# 1. Visualize Hi-C contact maps of ChrX in Normal people across polyploid samples
# 2. Visualize Hi-C contact maps of ChrX in HCC patient polyploid samples
# 3. Differential Hi-C analysis of ChrX between normal and HCC samples

# 

################################################################################ 

#' ## 1. Visualize Hi-C contact maps in Normal people across polyploid samples

```{r}

# Load GENOVA package
library(GENOVA)

# Load Hi-C contact matrices for different ploidy levels
N3_2n <- load_contacts(
signal_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/N3_2_N7_8/hicpro-output/hic_results/matrix/N3_2n/iced/500000/N3_2n_500000_iced.matrix',
indices_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/N3_2_N7_8/hicpro-output/hic_results/matrix/N3_2n/raw/500000/N3_2n_500000_abs.bed',
sample_name = "N3_2n")

N3_4n <- load_contacts(
signal_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/N3_4n/hicpro-output/hic_results/matrix/N3_4n/iced/500000/N3_4n_500000_iced.matrix',
indices_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/N3_4n/hicpro-output/hic_results/matrix/N3_4n/raw/500000/N3_4n_500000_abs.bed',
sample_name = "N3_4n")

N3_8n <- load_contacts(
signal_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/N3_8n/hicpro-output/hic_results/matrix/N3_8n/iced/500000/N3_8n_500000_iced.matrix',
indices_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/N3_8n/hicpro-output/hic_results/matrix/N3_8n/raw/500000/N3_8n_500000_abs.bed',
sample_name = "N3_8n")

# =========================
# Plot individual Hi-C contact maps
# =========================

# Plot 2n sample
print("2n N3(Normal)")
hic_matrixplot(
N3_2n,
chrom = "chrX", start = 0, end = 155270560,
cut.off = 2000 )



# Plot 4n sample
print("4n N3(Normal)")
hic_matrixplot(
N3_4n,
chrom = "chrX", start = 0, end = 155270560,
cut.off = 2000 )


# Plot 8n sample
print("8n N3(Normal)")
hic_matrixplot(
N3_8n,
chrom = "chrX", start = 0, end = 155270560,
cut.off = 2000 )

# =========================
# Plot differential Hi-C maps (exp2 - exp1)
# =========================

# 4n vs 2n
print("4n vs 2n N3(Normal)")
hic_matrixplot(exp1 = N3_4n,
              exp2 = N3_2n,
              coplot = 'diff',
              chrom = 'chrX',
              start = 0,
              end = 155270560, # upper limit of contacts
              cut.off = 1000)


# 8n vs 2n
print("8n vs 2n N3(Normal)")
hic_matrixplot(exp1 = N3_8n,
              exp2 = N3_2n,
              coplot = 'diff',
              chrom = 'chrX',
              start = 0,
              end = 155270560, # upper limit of contacts
              cut.off = 1000)

# 8n vs 4n
print("8n vs 4n N3(Normal)")
hic_matrixplot(exp1 = N3_8n,
              exp2 = N3_4n,
              coplot = 'diff',
              chrom = 'chrX',
              start = 0,
              end = 155270560, # upper limit of contacts
              cut.off = 1000)



```

#' ## 2. Visualize Hi-C contact maps in HCC patient polyploid samples

```{r}

# Load GENOVA package
library(GENOVA)

# Load Hi-C contact matrices for HCC patient with different ploidy levels
p13_2n <- load_contacts(
signal_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/P13_2n/hicpro-output/hic_results/matrix/P13_2n/iced/500000/P13_2n_500000_iced.matrix',
indices_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/P13_2n/hicpro-output/hic_results/matrix/P13_2n/raw/500000/P13_2n_500000_abs.bed',
sample_name = "p13_2n")

p13_4n <- load_contacts(
signal_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/P13_4n/hicpro-output/hic_results/matrix/P13_4n/iced/500000/P13_4n_500000_iced.matrix',
indices_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/P13_4n/hicpro-output/hic_results/matrix/P13_4n/raw/500000/P13_4n_500000_abs.bed',
sample_name = "p13_4n")

p13_8n <- load_contacts(
signal_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/P13_8n/hicpro-output/hic_results/matrix/P13_8n/iced/500000/P13_8n_500000_iced.matrix',
indices_path = '/public/home/mjy/HCC/Hi-C/polyploid/250508_P13_N3_N7_20M/P13_8n/hicpro-output/hic_results/matrix/P13_8n/raw/500000/P13_8n_500000_abs.bed',
sample_name = "p13_8n")

# =========================
# Plot individual Hi-C contact maps
# =========================

# Plot 2n sample
print("2n P13 (HCC)")
hic_matrixplot(
p13_2n,
chrom = "chrX", start = 0, end = 155270560,
cut.off = 2000)

# Plot 4n sample
print("4n P13 (HCC)")
hic_matrixplot(
p13_4n,
chrom = "chrX", start = 0, end = 155270560,
cut.off = 2000 )

# Plot 8n sample
print("8n P13 (HCC)")
hic_matrixplot(
p13_8n,
chrom = "chrX", start = 0, end = 155270560,
cut.off = 2000 )

```

#' ## 3. Differential Hi-C analysis of ChrX between normal and HCC samples

```{r}

# =========================
# Differential Hi-C maps (Normal vs HCC)
# =========================

# 2n comparison: N3 (normal) vs P13 (HCC)
print("2n comparison: N3 (normal) vs P13 (HCC)")
hic_matrixplot(exp1 = N3_2n,
              exp2 = p13_2n,
              coplot = 'diff',
              chrom = 'chrX',
              start = 0,
              end = 155270560,
              cut.off = 1000)



# 4n comparison: N3 (normal) vs P13 (HCC)
print("4n comparison: N3 (normal) vs P13 (HCC)")
hic_matrixplot(exp1 = N3_4n,
              exp2 = p13_4n,
              coplot = 'diff',
              chrom = 'chrX',
              start = 0,
              end = 155270560, # upper limit of contacts
              cut.off = 1000)



# 8n comparison: N3 (normal) vs P13 (HCC)
print("8n comparison: N3 (normal) vs P13 (HCC)")
hic_matrixplot(exp1 = N3_8n,
              exp2 = p13_8n,
              coplot = 'diff',
              chrom = 'chrX',
              start = 0,
              end = 155270560, # upper limit of contacts
              cut.off = 1000)



```

